---
title: "general visualization and data manipulation"
author: "Justin Herman"
date: "January 11, 2019"
output:
  html_document:
    theme: "simplex"
    highlight: 'pygments'
    toc: true
    toc_float: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r cars, message=FALSE}
library(tidyverse)
library(kableExtra)
library(knitr)
library(stringr)
library(readr)
library(plyr)
library(dplyr)
library(data.table)
library(readxl)
require(xlsx)
library(rJava)
library(xlsx)

```

# Tidy dataset 1: Movies database

+ [link to source](https://docs.google.com/spreadsheets/d/1K_a87dxLQR91GwVcqkqZK9kjxcYDl1tbK9EX9WvC7bo/edit#gid=601448852)

## Load in Data from xlsx file
+ Dataframe requires tidying
    + Join individual movies into one movies column
    + Separate timestamp by day and time
        + Could perform some interesting time series analysis on the effects of time of day of review
    + Omit NA Values
    
    
```{r}

movies <- read.xlsx("Movie_Ratings.xlsx",1)
movies%<>% 
    gather(Movies,n,3:12) %>% 
    select(Timestamp,Name,Movies,n,Number.of.Movies.Seen) %>%   
    separate('Timestamp',c("day","time")," ") %>% 
    na.omit() %>% 
    arrange(Name)
View(movies)
    
movies %>% 
    dplyr::group_by(Movies) %>% 
    dplyr::summarise(round(mean(n),2))


```

## data Visualization
+ Display average rating by user
+Display average rating by movie
```{r}


movies %>%  
    dplyr::group_by(Name) %>% 
    dplyr::summarise(Average_Rating_by_user=round(mean(n),2)) %>% 
    arrange(desc(Average_Rating_by_user)) %>% 
    ggplot(., aes(x=Name , Average_Rating_by_user)) + 
    geom_bar(aes(fill = Name), position = "dodge", stat = "identity")+
    theme(axis.text.x=element_text(angle=90,hjust=1))+
    coord_flip()

movies %>%  
    dplyr::group_by(Movies) %>% 
    dplyr::summarise(Average_movie_rating=round(mean(n),2)) %>% 
    arrange(desc(Average_movie_rating)) %>% 
    ggplot(., aes(x=Movies , Average_movie_rating)) + 
    geom_bar(aes(fill = Movies), position = "dodge", stat = "identity")+
    theme(axis.text.x=element_text(angle=45,hjust=1))
    



 

```

# World Bank info for Infant Mortality, Tuberculosis, and Aids
+ Binish recommended this dataset website
+ [link to source](http://apps.who.int/gho/data/node.country.country)
+ Randomly chose 4 countires with the letter `U`
+ Read in files individually/with functional
```{r}

#Manually Load in CSV
Uganda<- read.csv("https://raw.githubusercontent.com/justinherman42/Justin-Data-607/master/tidy_data_week4/uganda.csv",na.strings=c("","NA"))
Ukraine <- read.csv("https://raw.githubusercontent.com/justinherman42/Justin-Data-607/master/tidy_data_week4/Ukraine.csv",na.strings=c("","NA"))
United_kingdom <- read.csv("https://raw.githubusercontent.com/justinherman42/Justin-Data-607/master/tidy_data_week4/United_Kingdom.csv",na.strings=c("","NA"))
United_states <- read.csv("https://raw.githubusercontent.com/justinherman42/Justin-Data-607/master/tidy_data_week4/United_states.csv",na.strings=c("","NA"))


# Functional to load csv
files <- c("https://raw.githubusercontent.com/justinherman42/Justin-Data-607/master/tidy_data_week4/uganda.csv", "https://raw.githubusercontent.com/justinherman42/Justin-Data-607/master/tidy_data_week4/Ukraine.csv", "https://raw.githubusercontent.com/justinherman42/Justin-Data-607/master/tidy_data_week4/United_Kingdom.csv", "https://raw.githubusercontent.com/justinherman42/Justin-Data-607/master/tidy_data_week4/United_states.csv")
temp = files
myfiles <-lapply(temp,read.csv)
```


## Join on shared row index values(survey questions)
+ This will decrease our sample to only those questions that are answered in every dataset
```{r,warning= FALSE, message=FALSE}

merged_df<- inner_join(United_states,United_kingdom) %>% 
    inner_join(.,Ukraine) %>% 
    inner_join(.,Uganda)

# With Functional data
merged_df_2<- inner_join(myfiles[[1]],myfiles[[2]]) %>% 
    inner_join(myfiles[[3]]) %>%   
    inner_join(myfiles[[4]])
#kable(merged_df_2[1:15,])
```

## Tidy Dataset
```{r}

getting_tidy <- merged_df[-(1),] %>% 
    gather(United_states,Values,2:57) %>% 
    mutate(Year=rep(unlist(lapply(2015:2002,function(x) rep(x,23))),4)) %>% 
    spread(X, Values) 
```    

### Get rid of special characters ``]`` ``[``  ``-``   and add spacing for next pipe of tidying
+  Use pipe to Tidy Mean[min-max] columns into mean,min,max columns
```{r}

getting_tidy <-  as_data_frame(lapply(getting_tidy,function(x){ str_replace_all(x,"\\]|\\["," ")}))
getting_tidy <-  as_data_frame(lapply(getting_tidy,function(x){ str_replace_all(x,"-","  ")}))



final_df <- getting_tidy %>% 
    separate(.,`Incidence of tuberculosis (per 100 000 population per year)`,into=c('Mean_Incidence_Tuberculosis_100,000','Min_Tuberculosis',"Max_Tuberculosis")) %>% 
    separate(.,`Antiretroviral therapy coverage among people with HIV infection eligible for ART according to 2010 guidelines (%)`,into=c('Mean_Antiretroviral_coverage','Min_Antiretroviral_coverage','Max_Antiretroviral_coverage')) %>% 
    separate(.,`Infant mortality rate (probability of dying between birth and age 1 per 1000 live births)`,into=c('Mean_Infant mortality_Rate/1000','Min_Infant_Mortality rate',"Max_Infant_Mortality_Rate"),sep= "  ") %>% 
    separate(.,`Under-five mortality rate (probability of dying by age 5 per 1000 live births)`,into=c('Mean_Under-five_Mortality_Rate/1000','Min_Under-five_Mortality_Rate',"Max_Under-five_Mortality_Rate"), sep="  ") %>% 
    separate(.,`Tuberculosis treatment coverage`,into=c('Mean_Tuberculosis_Coverage','Min_Tuberculosis_Coverage',"Max_Tuberculosis_Coverage"))


```


###  More piping
+ First column needs to be renamed to country column 
+ Eliminate the .integers after the values in country column
+ Trim df
+ Convert columns to numeric for analysis

```{r}
final_df %<>% 
    plyr::rename(.,c('United_states'= 'Country')) %>% 
    mutate(Country=rep(c("Uganda","Ukraine","United_Kingdom","United_States"),each=14)) %>% 
    select(colnames(.)[c(1,2,10,11,12,13,14,15,16,17,21,22,23,30,31,32,33,34,35)]) %>% 
    arrange(.,`Country`)  

final_df[] <- lapply(final_df,function(x){ str_trim(x)}) 
final_df[3:19] <- lapply(final_df[3:19],function(x) as.numeric(as.character(x))) 
final_df <- as_data_frame(final_df)
kable(final_df)
```  

## Visual exploratory analysis of Infant Mortality
```{r}
final_df %>%
    mutate(GNP=as.numeric(`Gross national income per capita (PPP int. $)`)) %>% 
    dplyr::group_by(Country) %>% 
    ggplot(., aes(x=Year ,y=GNP))+ 
    geom_bar(aes(fill = Country), position = "dodge", stat = "identity")+
    theme(axis.text.x=element_text(angle=45,hjust=1))
    
final_df %>%
    dplyr::group_by(Country) %>% 
    ggplot(., aes(x=Year ,y=`Mean_Infant mortality_Rate/1000`))+ 
    geom_bar(aes(fill = Country), position = "dodge", stat = "identity")+
    theme(axis.text.x=element_text(angle=45,hjust=1))

final_df_display <- final_df %>% 
    arrange(Country,Year) %>% 
    select(.,Country,`Mean_Infant mortality_Rate/1000`)

```

+ Uganda seems to have reduced it's infant mortality rate by over 50%

## Numerical display of Infant Mortality
```{r}
## Numerical display     
kable(final_df_display)
```

## Display of infant Mortality with the 3 samples closer in scale

```{r}
final_df %>%
    dplyr::filter(., Country %in% c("United_Kingdom", "Ukraine","United_States")) %>% 
    dplyr::group_by(Country) %>% 
    ggplot(., aes(x=Year ,y=`Mean_Infant mortality_Rate/1000`))+ 
    geom_bar(aes(fill = Country), position = "dodge", stat = "identity")+
    theme(axis.text.x=element_text(angle=45,hjust=1))

```

+ Ukraine has seen a reduction of around 40%
+ US and UK have also seen a reduction, however UK seems to have a more substantial reduction(30%) over the US reduction of (17%)


## Look at Tuberculosis

```{r}

final_df %>%
    dplyr::filter(., Country %in% c("Uganda", "Ukraine","United_Kingdom","United_States")) %>% 
    dplyr::group_by(Country) %>% 
    ggplot(., aes(x=Year ,y=`Mean_Incidence_Tuberculosis_100,000`))+ 
    geom_bar(aes(fill = Country), position = "dodge", stat = "identity")+
    theme(axis.text.x=element_text(angle=45,hjust=1))
```

+ Uganda seems to have seen some moderate improvement over the past 13 years in tuberculosis, but the rate is relatively high still
+ Ukraine seems to have has a similar improvement

## Look at Tuberculosis in Heavily industrialized nations( UK, US)

```{r}
final_df %>%
    dplyr::filter(., Country %in% c("United_Kingdom","United_States")) %>% 
    dplyr::group_by(Country) %>% 
    ggplot(., aes(x=Year ,y=`Mean_Incidence_Tuberculosis_100,000`))+ 
    geom_bar(aes(fill = Country), position = "dodge", stat = "identity")+
    theme(axis.text.x=element_text(angle=45,hjust=1))
```

+ England appears to have made some moderate improvement in the past several years
+ Tuberculosis in the United states is almost non-existant and it appears it is decreasing




# Third Dataset- Gapminder
+ Burcu recommended the Gapminder data website
+ [link to source](https://www.gapminder.org/data/)
+ mix of WHO datasets and 

## War or food?
+ Are nations dealing with war more food insecure?
+ How do state expenditures differ on war versus food?

## Load in third dataset 
+ file 1 iis a file with death rates out of 100k from war  [link to source](https://docs.google.com/spreadsheets/d/15kseoFj-w39TvMl_Ije-C2TZy-DKYyECsF-uGdec0z8/pub)
+ file 2 is a measurement of daily calories consumption    [link to source](https://docs.google.com/spreadsheets/d/14G6CjF6NblTGf6kkQclpXp3XZ3D4Nkw1I92DB4fOjXo/pub)
+file 3 is GDP War expenditure %                           [link to source](https://docs.google.com/spreadsheets/d/1HXH3ncdQLFpOXVNdC2nkPCSGz_perkfPcBeU1e2HF24/pub)
+ file 4 is GDP Agricultural Expenditure                   [link to source](https://docs.google.com/spreadsheets/d/16JAXDjaKcyaEfd6sO_pRC892NG8VAk6moyw3Gl50-Ys/pub)

```{r}
delete.na <- function(DF, n=0) {
  DF[rowSums(is.na(DF)) <= n,]
}


War<- read.xlsx("War.xlsx", header=TRUE,1,)
food_security <- read.xlsx("indicator food_consumption.xlsx", header=TRUE,1,)
Agricultural_expend<- read.xlsx("Agricultural_expenditure.xlsx", header=TRUE,1,)
military_expend <- read.xlsx("military_expenditure.xlsx", header=TRUE,1,)

dim(food_security)
dim(military_expend)
dim(Agricultural_expend)
dim(War)
```

## DF Preparation 
+ Eliminate rows with substantial NA's
+ Convert columns to numeric 
+ Rename columns to acoid duplication of Column name sfor eventual Join
+ Filter df's for common shared year range of 1988-2007
```{r}
military_expend <- delete.na(military_expend, 20)
Agricultural_expend <- delete.na(Agricultural_expend, 15)
food_security <- delete.na(food_security, 20)
War <- as_data_frame(lapply(War, function(x) if(is.numeric(x)) round(x,2) else x))
military_expend <- as_data_frame(lapply(military_expend, function(x) if(is.numeric(x)) round(x,2) else x))
War <- War[2:4]
colnames(War) <- c("Country", 'war_2002','war_2004')
colnames(military_expend) <-  c("Country",paste(1988:2011,'Military_expend'))
colnames(food_security) <- c("Country",paste (1961:2007,"food_security"))
colnames(Agricultural_expend) <- c("Country",paste(1961:2011,"Agricultural_expend"))
Agricultural_expend <- Agricultural_expend[c(1,29:48)]
food_security <- food_security[c(1,29:48)]
military_expend <- military_expend[c(1:21)]


merged_df_3<- inner_join(War,military_expend) %>%
    inner_join(.,food_security) %>% 
    inner_join(.,Agricultural_expend)  
dim(merged_df_3)

```


## Exploratory Analysis of Death from wars data 
+ explore war death df
```{r}
merged_df_3$war_2002-merged_df_3$war_2004


merged_df_3[1:3] %>% 
    mutate(.,average_deaths=(war_2002+war_2004)/2) %>% 
    arrange(.,average_deaths) %>%  
    kable(.)
 
```

## Create bins of death from wars to categorize Nations 
+ Add average deaths from war column
+ Categorize deaths from war by levels
    + no deaths, low_deaths, moderate_deaths, high_deaths

```{r}
food_war <- merged_df_3 %>% 
    gather(Wars,Values,4:63) %>% 
    separate(.,Wars,into= c("Year", "Type"),sep=" ") %>% 
    spread(Type,Values) %>% 
    mutate(.,average_deaths=(war_2002+war_2004)/2) %>% 
    select(-c(war_2002,war_2004)) %>% 
    mutate(Deaths_from_war=cut(average_deaths, breaks=c(-1,0,1,10,Inf), labels= c("no_deaths", "low_deaths", "moderate_deaths", "high_deaths")))
   
    

kable(food_war[1:50,])
```

## Summary Explorations
```{r}
# 
food_war <-food_war %>%
    dplyr::group_by(Deaths_from_war,Year)
agricultural_spending_by_year <- food_war %>%
    dplyr::summarize(agri_exp_mean=mean(Agricultural_expend, na.rm = TRUE))
#food_war %>%
   # dplyr::summarize(mean(Military_expend, na.rm = TRUE))
food_security_by_year <- food_war %>%
    dplyr::summarize(daily_calories=mean(food_security, na.rm = TRUE))

food_war <-food_war %>% 
     dplyr::group_by(Deaths_from_war)
# food_war %>% 
#     dplyr::summarize(mean(Agricultural_expend, na.rm = TRUE))
# food_war %>% 
#     dplyr::summarize(mean(Military_expend, na.rm = TRUE))
# food_war %>% 
#     dplyr::summarize(mean(food_security, na.rm = TRUE))
# 
# food_war    
```

## Observations
+ My categories aren't of even sizes so my clustering could effect my observations and lead to large impact from outliers
+ I don't have an overall GDP per country, this could make a big difference in relative % GDP expenditures

+Tentative Observations
    + Areas with the most death from wars seem to spend more of their GDP on agricultural expenditures
    + Areas with the most deaths don't spend much on military expenditures, and low death areas actually spend more than high death areas 
    + High and moderate death areas are significantly more food insecure. Calorie consumption is about 30% lower in these areas




## Data Visualization 
```{r}
food_security_by_year %>% 
    ggplot(.,aes(x=Year,y=daily_calories))+
    geom_bar(aes(fill = Deaths_from_war), position = "dodge", stat = "identity")+
    theme(axis.text.x=element_text(angle=45,hjust=1))

```

+ Caloric consumption seems rather Flat in all the categories over the last three decades


```{r}
agricultural_spending_by_year %>% 
    ggplot(.,aes(x=Year,y=agri_exp_mean))+
    geom_bar(aes(fill = Deaths_from_war), position = "dodge", stat = "identity")+
    theme(axis.text.x=element_text(angle=45,hjust=1))
```

+ As a percent of GDP agricultural expenditure has decreased across all categories over last three decades
+ Area's with higher casualties of war seem to have made large cuts to expenditures

```{r}
food_war %>% 
    ggplot(.,aes(x=Agricultural_expend,y=food_security,shape=Deaths_from_war,color=Deaths_from_war))+
    geom_point()
food_war %>% 
    filter(.,Military_expend <20) %>% 
    ggplot(.,aes(x=Military_expend,y=food_security,shape=Deaths_from_war,color=Deaths_from_war))+
    geom_point()
```


## The  Scatter plots are very Interesting 
+ A handful of outliers were removed for better visualization 
### Graph 1 Agricultural exp vs Food Security
+ Two Clear Clusters develop in the first graph
    + Areas with high death rates seem to be the least food secure and silumtaneously spend the most on agricultural subsidy
    + Areas with low death rates spends the least on food subsidy, and are the most food secure

### Graph 2 Military exp vs Food Security 
+ Areas with low death rates seem to spend the most on the Military


